---
title: "Assignment 6"
author: "Kaodinna Onwumechili"
format: pdf
---

```{r}
suppressWarnings({
  library(DBI)
  library(duckdb)
  library(nycflights13)
  library(dbplyr)
})
```

## Part I: Answer the following question by running SQL using RStudio.

#### a. List the first 10 rows of the `flights` table.

```{r}
dbconn <- dbConnect(duckdb())
dbWriteTable(dbconn, "flights", flights)
```

```{r}
dbGetQuery(
  dbconn,
  # select all columns
  # from flights
  # only first 10 rows
  "SELECT *
  FROM flights
  LIMIT 10;"
)
```

#### b. Count the Number of Flights for Each Carrier

Find the **number of flights** for each carrier.

    - Your output should include:
        - carrier
        - number of flights
        
Sort from the **largest to smallest**.

```{r}
dbGetQuery(
  dbconn,
  "SELECT carrier,
  COUNT(*) AS num_flights
  FROM flights
  GROUP BY carrier
  ORDER BY num_flights DESC;"
)
```

#### c. Find the **average departure delay by month**

    - Your output should include:
        - month
        - average departure delay
        
```{r}
dbGetQuery(
  dbconn,
  "SELECT month, AVG(dep_delay) AS avg_dep_delay
  FROM flights
  GROUP BY month;"
)
```

