---
title: "Assignment 6"
author: "Kaodinna Onwumechili"
format: pdf
---

```{r}
suppressWarnings({
  library(DBI)
  library(duckdb)
  library(nycflights13)
  library(dbplyr)
})
```

## Part I: Answer the following question by running SQL using RStudio.

#### a. List the first 10 rows of the `flights` table.

```{r}
dbconn <- dbConnect(duckdb())
dbWriteTable(dbconn, "flights", flights)
```

```{r}
dbGetQuery(
  dbconn,
  # select all columns
  # from flights
  # only first 10 rows
  "SELECT *
  FROM flights
  LIMIT 10;"
)
```

#### b. Find the **number of flights**

###### i. in the dataset

```{r}
# count all columns, only show the count column
# rename it as num_flights for increased readability
dbGetQuery(dbconn, "SELECT COUNT(*) AS num_flights FROM flights;")
```

###### ii. for each carrier

```{r}
# select and group by carrier, as we are counting the number of flights for each
# carrier
dbGetQuery(
  dbconn,
  "SELECT carrier, COUNT(*) AS num_flights
  FROM flights
  GROUP BY carrier
  ORDER BY num_flights DESC;"
)
```

#### c. Find the **average departure delay** (`dep_delay`)

```{r}
dbGetQuery(dbconn, "SELECT AVG(dep_delay) AS avg_dep_delay FROM flights;")
```

#### d. Find the **top 5 destinations** (`dest`) with the most flights.

    - Your output should be:
        - destination
        - number of flights
        
Sort from the **largest to smallest**.

```{r}
dbGetQuery(
  dbconn,
  "SELECT dest, COUNT(*) as num_flights
  FROM flights
  GROUP BY dest
  ORDER BY num_flights DESC
  LIMIT 5;" # LIMIT shows the first n rows
)
```

#### e. Create a table showing:

    - `carrier`
    - average arrival delay (`arr_delay`)
    
Sort the results from **largest to smallest delay**.

```{r}
dbGetQuery(
  dbconn,
  "SELECT carrier, AVG(arr_delay) AS avg_arr_delay
  FROM flights
  GROUP BY carrier
  ORDER BY avg_arr_delay DESC;"
)

dbDisconnect(dbconn) # have to disconnect when done
```

